---
title: 'Peer-graded Assignment: Course Project 1'
author: "Maulid Hussein Bwabo"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
# Reproducible Research: Peer Assessment 1
```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
install.packages("Hmisc")
library(Hmisc)
```
## Loading and preprocessing the data
### 1. Load the data (i.e. read.csv())
```{r, results='markup', warning=TRUE, message=TRUE}
library(readr)
activity = read_csv("C:/Users/bwabo/OneDrive/Desktop/Data Science/repdata_data_activity/activity.csv")
View(activity)

`##### 2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
#activity$interval = strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activity$interval), format='%H:%M')
```
## What is mean total number of steps taken per day?
```{r}
stepsByDay = tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```
##### 1. Make a histogram of the total number of steps taken each day
```{r}
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```
##### 2. To calculate and report the mean and median total number of steps taken per day
```{r}
stepsByDayMean = mean(stepsByDay)
stepsByDayMedian = median(stepsByDay)
```
* Mean: `r stepsByDayMean`
* Median:  `r stepsByDayMedian`

-----
## What is the average daily activity pattern?
```{r}
averageStepsPerTimeBlock = aggregate(x=list(meanSteps=activity$steps), by=list(interval=activity$interval), FUN=mean, na.rm=TRUE)
```
##### 1. Making a time series plot
```{r}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```
##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
mostSteps = which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps =  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```
* Most Steps at: `r timeMostSteps`
----
## Imputing missing values from the activity data set
##### 1. Calculate and report the total number of missing values in the data set 
```{r}
numMissingValues = length(which(is.na(activity$steps)))
```
##### 2. Numerous strategy to handle the missing values in the activity dataset.
##### 3. Creating new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activityImputed = activity
activityImputed$steps = impute(activity$steps, fun=mean)
```
##### 4. Making an histogram of the total number of steps taken each day 
```{r}
stepsByDayImputed = tapply(activityImputed$steps, activityImputed$date, sum)
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```
##### ... as well as Calculating and reporting the mean and median total number of steps taken per day. 
```{r}
stepsByDayMeanImputed = mean(stepsByDayImputed)
stepsByDayMedianImputed = median(stepsByDayImputed)
```
* Mean (Imputed): `r stepsByDayMeanImputed`
* Median (Imputed):  `r stepsByDayMedianImputed`
## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
activityImputed$dateType <-  ifelse(as.POSIXlt(activityImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```
##### 2. Producing panel plot that containing a time series 

```{r}
averagedActivityImputed = aggregate(steps ~ interval + date
                        ,data=activityDataImputed, mean)
ggplot(averagedActivityImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(date ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```