---
title: 'Peer-graded Assignment: Course Project 1'
author: "Maulid Hussein Bwabo"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
## Loading and preprocessing the data
### 1. Load the data (i.e. read.csv())
```{r, results='markup', warning=TRUE, message=TRUE}
activity1=read.csv("C:/Users/bwabo/OneDrive/Desktop/Data Science/repdata_data_activity/activity.csv")
```
`##### 2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
#activity1$interval = strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activity1$interval), format='%H:%M')
```
## What is mean total number of steps taken per day?
```{r}
stepsByDay1 = tapply(activity1$steps, activity1$date, sum, na.rm=TRUE)
```
##### 1. Make a histogram of the total number of steps taken each day
```{r}
ggplot2::qplot(stepsByDay1, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```
##### 2. To calculate and report the mean and median total number of steps taken per day
```{r}
stepsByDay1Mean = mean(stepsByDay1)
stepsByDay1Median = median(stepsByDay1)
```

-----
## What is the average daily activity pattern?
```{r}
averageStepsPerTimeBlock1 = aggregate(x=list(meanSteps=activity1$steps), by=list(interval=activity1$interval), FUN=mean, na.rm=TRUE)
```
##### 1. Making a time series plot
```{r}
ggplot2::ggplot(data=averageStepsPerTimeBlock1, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```
##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
mostSteps1 = which.max(averageStepsPerTimeBlock1$meanSteps)
timeMostSteps =  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock1[mostSteps,'interval'])
```
* Most Steps at: `r timeMostSteps`
----
## Imputing missing values from the activity data set
##### 1. Calculate and report the total number of missing values in the data set 
```{r}
numMissingValues1 = length(which(is.na(activity1$steps)))
```
##### 2. Numerous strategy to handle the missing values in the activity dataset.
##### 3. Creating new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activity1Imputed = activity1
activity1Imputed$steps = impute(activity1$steps, fun=mean)
```
##### 4. Making an histogram of the total number of steps taken each day 
```{r}
stepsByDayImputed1 = tapply(activity1Imputed$steps, activity1Imputed$date, sum)
qplot(stepsByDayImputed1, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```
##### ... as well as Calculating and reporting the mean and median total number of steps taken per day. 
```{r}
stepsByDayMeanImputed1 = mean(stepsByDayImputed1)
stepsByDayMedianImputed1 = median(stepsByDayImputed1)
```
* Mean (Imputed): `r stepsByDayMeanImputed`
* Median (Imputed):  `r stepsByDayMedianImputed`
## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
activity1Imputed$dateType =  ifelse(as.POSIXlt(activity1Imputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```
##### 2. Producing panel plot that containing a time series 

```{r}
averagedActivity1Imputed = aggregate(steps ~ interval + date
                        ,data=activityDataImputed, mean)
ggplot(averagedActivity1Imputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(date ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```